---
title: "Data Analysis"
author: "Alina Cherkas"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, fig.align='center')
```

# {.tabset .tabset-fade .tabset-pills}

## Introduction

The notebook can be used to replicate the results of the analysis in *The Effects of Migration on Attitudes towards the European Union: Extent, Dynamics and Causality*, master's thesis submitted to CEU in a.y. 2019/2020. In the analysis, two datasets are used:

1. `DatasetFinal` from `CEU_Thesis_Data.xlsx`, a panel country-level dataset.
2. `CEU_Thesis_Multilevel.feather`, a cross-section of multi-level data (individual, regional and country levels).

The first dataset **can be replicated** in its whole using `1.1. Data Preparation Country.ipynb`.
Due to data sharing restrictions, individual-level records used to construct the second dataset are not provided. Thus, the second dataset **cannot be replicated**, but the code used to build it can be found in `1.2. Data Preparation Multilevel.ipynb`. The script also contains a link to the original data source, from which the records can be requested.

The below code can be run step-by-step to replicate the main results of the study. However, the code in sections `Multilevel Data`, `Table 6` and `Table 7` is only for demonstration purposes, since no data are provided for replication as described above.

## Prerequisites

**Note:** you might need to install some of the packages below befor running the code.

```{r Libraries, class.source='fold-show'}
# Standard Imports
library(tidyverse)
library(magrittr)

# Input/Output
library(readxl)
library(feather)

# For tables and graphs
library(stargazer)
library(ggdag)
library(dagitty)

# For Panel Models
library(plm)

# For Clustered Stanrard Errors
library(clubSandwich)

# For multilevel models
library(lme4)
```


## Dataset

```{r, Country Data, class.source='fold-show'}
df.Main <- read_excel("CEU_Thesis_Data.xlsx", sheet = "DatasetFinal")
df.Main %<>% pdata.frame(index = c("Country", "Year"))
```

```{r, Full Sample Description}
df_lambda <- df.Main %>% select(Country, Year, ImmigrantStock, InflowsOECD, InflowsEurostat, ShiftShare,
                                 Positive, Negative, VeryNegative)

get_mean <- function(x){mean(x,na.rm=TRUE) %>% round(3)}

DT::datatable(df_lambda %>% select(-Year) %>% group_by(Country) %>% summarise_each(get_mean),
              caption = "Table X. Mean Country-level Values of Main Variables in the Full Dataset (2009-2017)",
              extensions = c('FixedColumns',"FixedHeader")
              )

DT::datatable(df_lambda %>% select(-Country) %>% group_by(Year) %>% summarise_each(get_mean),
              caption = "Table X. Mean Yearly Values of Main Variables in the Full Dataset (2009-2017)",
              extensions = c('FixedColumns',"FixedHeader")
              )

DT::datatable(df_lambda %>% filter(Country != "Luxembourg") %>% select(-Country) %>% group_by(Year) %>%
                summarise_each(get_mean),
              caption = "Table X. Mean Yearly Values of Main Variables in the Full Dataset without Luxembourg (2009-2017)",
              extensions = c('FixedColumns',"FixedHeader")
              )
```

```{r, IV Sample Description}
df_lambda <- df.Main %>% filter(IV_Sample == TRUE) %>% 
  select(Country, Year, ImmigrantStock, InflowsOECD, InflowsEurostat, ShiftShare, Positive, Negative, VeryNegative)

DT::datatable(df_lambda %>% select(-Year) %>% group_by(Country) %>% summarise_each(get_mean),
              caption = "Table X. Mean Country-level Values of Main Variables in the IV Subset (2009-2017)",
              extensions = c('FixedColumns',"FixedHeader")
              )

DT::datatable(df_lambda %>% select(-Country) %>% group_by(Year) %>% summarise_each(get_mean),
              caption = "Table X. Mean Yearly Values of Main Variables in the IV Subset (2009-2017)",
              extensions = c('FixedColumns',"FixedHeader")
              )
```

## Table 1

Replication can be found in `1.1. Data Preparation Country.ipynb`

## Figure 1

The figure was created using _Tableau Desktop (2020.2)_ based on `DatasetFinal` from `CEU_Thesis_Data.xlsx`. Replication codes are not applicable.

## Figure 2

The figure was created using _Tableau Desktop (2020.2)_ based on `DatasetFinal` from `CEU_Thesis_Data.xlsx`. Replication codes are not applicable.

## Figure 3

```{r, fig.cap='Figure 3. Directed Graph of the Causal Relationship between Immigration and Attitudes'}
g <- dagify(y_it ~ c_it + w_1it + w_2it + w_3it,
            y_it ~~ x_it,
            x_it ~ c_it + z_it)

to_labels <- c(expression(c[it]),
               expression(w[1][it]),
               expression(w[2][it]),
               expression(w[3][it]),
               expression(x[it]),
               expression(z[it]),
               expression(y[it]))

g %>% tidy_dagitty() %>% dag_label(labels = c("y_it" = "Attitudes to the EU",
                                              "x_it" = "Immigration",
                                              "c_it" = "Omitted Variable",
                                              "z_it" = "Past Settlement Instrument",
                                              "w_1it" = "Unemployment",
                                              "w_2it" = "GDP Growth",
                                              "w_3it" = "Pop. Age Composition"
                                              )) %>% 
  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_edges() + 
  geom_dag_node(color = "#2c3e50", alpha = 0.8) +
  geom_dag_text(color = "white",
                parse = TRUE,
                label = to_labels) +
  geom_dag_label_repel(aes(label = label),
                       fill = "#a8e6cf",
                       alpha = 0.8,
                       nudge_x = .3,
                       nudge_y = .3) +
  theme_void() + theme()
```

## Figure 4

Replication can be found in `1.1. Data Preparation Country.ipynb`.

## Table 2

```{r, Baseline Unit Root Testing}
purtest(InflowsOECD ~ 1, data = df.Main, index = "Country", lag = 0, test = "levinlin")
purtest(InflowsOECD ~ 1, data = df.Main, index = "Country", lag = 1, test = "levinlin")
purtest(InflowsOECD ~ 1, data = df.Main, index = "Country", lag = 0, test = "ips")
purtest(InflowsOECD ~ 1, data = df.Main, index = "Country", lag = 1, test = "ips")
```

```{r, Baseline Pooled Models, class.source='fold-show'}
# Complete Pooling
m.OLS1 <- plm(VeryNegative ~ InflowsOECD + ImmigrantStock + UnempRate, data = df.Main, model = "pooling")
m.OLS2 <- update(m.OLS1, .~. + GlobalExposure + GDP_Growth)
m.OLS3 <- update(m.OLS2, .~. + EduShare + Pop25_49 +  PopOver65)

# Clustered Standard Errors
m.OLS1_se <- coef_test(m.OLS1, vcov = "CR1", cluster = "individual")
m.OLS2_se <- coef_test(m.OLS2, vcov = "CR1", cluster = "individual")
m.OLS3_se <- coef_test(m.OLS3, vcov = "CR1", cluster = "individual")
```

```{r, Baseline FE Models, class.source='fold-show'}
# Fixed-Effects
m.FE1 <- update(m.OLS3, model = "within", effect = "twoways")
m.FE2 <- update(m.FE1, .~. - InflowsOECD + lag(InflowsOECD,1))
m.FE3 <- update(m.FE1, .~. + lag(InflowsOECD,1))

# Clustered Standard Errors
m.FE1_se <- coef_test(m.FE1, vcov = "CR1", cluster = "individual")
m.FE2_se <- coef_test(m.FE2, vcov = "CR1", cluster = "individual")
m.FE3_se <- coef_test(m.FE3, vcov = "CR1", cluster = "individual")
```

```{r, Baseline Results, results='asis'}
stargazer(m.OLS1, m.OLS2, m.OLS3, m.FE1, m.FE2, m.FE3, type = "html", style = "apsr",
          se = list(m.OLS1_se$SE, m.OLS2_se$SE, m.OLS3_se$SE, m.FE1_se$SE, m.FE2_se$SE, m.FE3_se$SE),
          title = "Table 2. Baseline Results for 22 European Countries (2009-2017)",
          column.labels =  c("Complete Pooling", "Fixed-Effects"),
          column.separate = c(3, 3),
          model.numbers = TRUE,
          dep.var.labels = "Share of Population with Very Negative Attitude towards the EU",
          covariate.labels = c("Immigrant Flows", 
                               "Immigrant Stock", 
                               "Unemployment Rate",
                               "Globalisation Exposure", 
                               "GDP Growth (in %)",
                               "Education (in %) ",
                               "Share Aged 25-49",
                               "Share Aged Over 65", 
                               "Immigrant Flows (1-year lag)")
          )
```

## Table 3

For first-stage results, see Annex A.

```{r, IV Sample, class.source='fold-show'}
df.IV <- df.Main %>% filter(IV_Sample == TRUE) %>% as_tibble()
df.IV %<>% pdata.frame(index = c("Country", "Year"))
```

```{r, IV Unit Root Testing}
purtest(ShiftShare ~ 1, data = df.IV, index = "Country", lag =0, test = "levinlin")
purtest(ShiftShare ~ 1, data = df.IV, index = "Country", lag =1, test = "levinlin")
purtest(ShiftShare ~ 1, data = df.IV, index = "Country", lag = 0, test = "ips")
purtest(ShiftShare ~ 1, data = df.IV, index = "Country", lag = 1, test = "ips")
```

```{r, IV Sample Models, class.source='fold-show'}
# Fixed Effects
m.FE4 <- plm(VeryNegative ~ InflowsOECD + ImmigrantStock + UnempRate + Pop25_49, data = df.IV, model = "within",
             effect = "twoways")
m.FE5 <- update(m.FE4, .~. + GDP_Growth + GlobalExposure)
m.FE6 <- update(m.FE5, .~. + lag(InflowsOECD,1))

# First Stage
m.1stage1 <- plm(InflowsOECD ~ ShiftShare + ImmigrantStock + UnempRate,
                 data = df.IV, model = "within", effect = "twoways")
m.1stage2 <- update(m.1stage1, .~. + Pop25_49 + GDP_Growth + GlobalExposure)

# IV Fixed Effects
m.IVFE1 <- update(m.FE4, .~. -InflowsOECD +  fitted(m.1stage1))
m.IVFE2 <- update(m.FE5, .~. -InflowsOECD + fitted(m.1stage2))
m.IVFE3 <- update(m.IVFE2, .~. + lag(fitted(m.1stage2),1))

# Clustered Standard Errors
m.FE4_SE <- coef_test(m.FE4, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
m.FE5_SE <- coef_test(m.FE5, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
m.FE6_SE <- coef_test(m.FE6, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
m.IVFE1_SE <- coef_test(m.IVFE1, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
m.IVFE2_SE <- coef_test(m.IVFE2, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
m.IVFE3_SE <- coef_test(m.IVFE3, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
```

```{r, IV Sample Results, results='asis'}
stargazer(m.FE4, m.FE5, m.FE6, m.IVFE1, m.IVFE2, m.IVFE3, type = "html", style = "apsr",
          se = list(m.FE4_SE$SE, m.FE5_SE$SE, m.FE6_SE$SE, m.IVFE1_SE$SE, m.IVFE2_SE$SE, m.IVFE3_SE$SE),
          title = "Table 3. Main Results for a Subset of 7 European Countries (2009-2017)",
          column.labels =  c("Fixed-Effects", "IV Fixed-Effects"),
          column.separate = c(3, 3),
          model.numbers = TRUE,
          dep.var.labels = "Share of Population with Very Negative Attitude towards the EU",
          covariate.labels = c("Immigrant Flows",
                               "Immigrant Stock",
                               "Unemployment Rate",
                               "Share Aged 25-49",
                               "GDP Growth (in %)",
                               "Globalisation Exposure",
                               "Immigrant Flows (1-year lag)",
                               "Immigrant Flows (IV)",
                               "Immigrant Flows (IV)",
                               "Immigrant Flows (IV, 1-year lag)")
          )
```

## Table 4

For first-stage results, see Annex A.

```{r, First Robustness Check, class.source='fold-show'}
# Fixed Effects
m.FE7 <- update(m.FE4, .~. - InflowsOECD + InflowsEurostat)
m.FE8 <- update(m.FE7, .~. + GDP_Growth + GlobalExposure)
m.FE9 <- update(m.FE8, .~. + lag(InflowsEurostat,1))

# First Stage
m.1stage3 <- update(m.1stage1, InflowsEurostat~. )
m.1stage4 <- update(m.1stage2, InflowsEurostat~. )

# IV Fixed Effects
m.IVFE4 <- update(m.FE7, .~. -InflowsEurostat +  fitted(m.1stage3))
m.IVFE5 <- update(m.FE8, .~. -InflowsEurostat + fitted(m.1stage4))
m.IVFE6 <- update(m.IVFE5, .~. -InflowsEurostat + lag(fitted(m.1stage4),1))

# Clustered Standard Errors
m.FE7_SE <- coef_test(m.FE7, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
m.FE8_SE <- coef_test(m.FE8, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
m.FE9_SE <- coef_test(m.FE9, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
m.IVFE4_SE <- coef_test(m.IVFE1, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
m.IVFE5_SE <- coef_test(m.IVFE2, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
m.IVFE6_SE <- coef_test(m.IVFE3, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
```

```{r, First Robustness Check Results, results='asis'}
stargazer(m.FE7, m.FE8, m.FE9, m.IVFE4, m.IVFE5, m.IVFE6, type = "html", style = "apsr",
          se = list(m.FE7_SE$SE, m.FE8_SE$SE, m.FE9_SE$SE, m.IVFE4_SE$SE, m.IVFE5_SE$SE, m.IVFE6_SE$SE),
          title = "Table 4. First Robustness Check of Main Results",
          column.labels =  c("Fixed-Effects", "IV Fixed-Effects"),
          column.separate = c(3, 3),
          model.numbers = TRUE,
          dep.var.labels = "Share of Population with Very Negative Attitude towards the EU",
          covariate.labels = c("Immigrant Stock",
                               "Unemployment Rate",
                               "Share Aged 25-49",
                               "Immigrant Flows (Eurostat)",
                               "GDP Growth (in %)",
                               "Globalisation Exposure",
                               "Immigrant Flows (1-year lag)",
                               "Immigrant Flows (Eurostat, IV)",
                               "Immigrant Flows (Eurostat, IV)",
                               "Immigrant Flows (Eurostat, IV, 1-year lag)")
          )
```

## Table 5

For first-stage results, see Annex A.

```{r, Second Robustness Check, class.source='fold-show'}
m.FE4 <- update(m.FE4, Negative~.)
m.FE5 <- update(m.FE4, Negative~.)
m.FE6 <- update(m.FE6, Negative~.)

m.IVFE1 <- update(m.IVFE1, Negative~.)
m.IVFE2 <- update(m.IVFE2, Negative~.)
m.IVFE3 <- update(m.IVFE3, Negative~.)

# Clustered Standard Errors
m.FE4_SE <- coef_test(m.FE4, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
m.FE5_SE <- coef_test(m.FE5, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
m.FE6_SE <- coef_test(m.FE6, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
m.IVFE1_SE <- coef_test(m.IVFE1, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
m.IVFE2_SE <- coef_test(m.IVFE2, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
m.IVFE3_SE <- coef_test(m.IVFE3, vcov = "CR1", cluster = "individual", test = "Satterthwaite")
```

```{r, Second Robustness Check Results, results='asis'}
stargazer(m.FE4, m.FE5, m.FE6, m.IVFE1, m.IVFE2, m.IVFE3, type = "html", style = "apsr",
          se = list(m.FE4_SE$SE, m.FE5_SE$SE, m.FE6_SE$SE, m.IVFE1_SE$SE, m.IVFE2_SE$SE, m.IVFE3_SE$SE),
          title = "Table 5. Second Robustness Check of Main Results)",
          column.labels =  c("Fixed-Effects", "IV Fixed-Effects"),
          column.separate = c(3, 3),
          model.numbers = TRUE,
          dep.var.labels = "Share of Population with Fairly or Very Negative Attitude towards the EU",
          covariate.labels = c("Immigrant Flows (OECD)",
                               "Immigrant Stock",
                               "Unemployment Rate",
                               "Share Aged 25-49",
                               "GDP Growth (in %)",
                               "Globalisation Exposure",
                               "Immigrant Flows (1-year lag)",
                               "Immigrant Flows (IV)",
                               "Immigrant Flows (IV)",
                               "Immigrant Flows (IV, 1-year lag)")
          )
```

## Multilevel Data

```{r, Multilevel Data, class.source='fold-show'}
df_multilevel <- read_feather('CEU_Thesis_Multilevel.feather')
df_multilevel %<>% mutate(Age = scale(Age))
df_multilevel %<>% drop_na()
summary(df_multilevel)
```

## Table 6

```{r, class.source='fold-show'}
m0 <- lm(Attitude ~ Gender + Age + Residence + VoiceCounts + LifeSatisfaction, data = df_multilevel)
m1 <- lm(Attitude ~ Gender + Age + Residence + VoiceCounts + LifeSatisfaction + factor(ISO) - 1, data = df_multilevel)
m2 <- lm(Attitude ~ Gender + Age + Residence + VoiceCounts + LifeSatisfaction + factor(NUTS) - 1,
         data = df_multilevel)
```

```{r, OLS Results, results='asis'}
stargazer(m0, m1, m2, type = "html", style = "apsr",
          title = "Table 6. Pooled Cross-sectional Model for a Subset of European Countries (2015)",
          column.labels =  c("Complete Pooling", "No Pooling"),
          column.separate = c(1,2), keep = 1:6,
          covariate.labels = c("Gender (Female = 1)", "Age (Scaled)", "Residence (Large City = 1)", 
                               "Voice Counts (Binary)", "Life Satisfaction (Binary)"),
          dep.var.labels = "Attitude Towards the EU (5-point scale)",
          add.lines = list(c("Country-fixed Effects", "No", "Yes", "No"),
                           c("NUTS-fixed Effects", "No", "No", "Yes")))
```

## Table 7

For variance components, see outputs below.

```{r, Multilevel Models, class.source='fold-show'}
ml0 <- lmer(Attitude ~ 1 + (1|NUTS)+ (1|ISO), data = df_multilevel)
ml1 <- update(ml0, .~. + Gender + Age + Residence + VoiceCounts + LifeSatisfaction)
ml2 <- update(ml1, .~. + StockTotal + InflowsOECD)
ml3 <- update(ml2, .~. + NativeEdu + NativeUnempl)
ml4 <- update(ml2, .~. - (1|ISO) + (1 + Gender + Age|ISO))
```

```{r, Results Multilevel, results='asis'}
stargazer(ml0, ml1, ml2, ml3, ml4, type = "html", style = "apsr",
          title = "Table 7. Multilevel Cross-sectional Model for a Subset of European Countries (2015)",
          column.labels =  c("Null Model", "Level 1", "Nested Model"),
          column.separate = c(1, 1, 3),
          covariate.labels = c("Gender (Female = 1)", "Age (Scaled)", "Residence (Large City = 1)", 
                               "Voice Counts (Binary)", "Life Satisfaction (Binary)",
                               "Immigant Stock (NUTS)", "Immigant Flow (Country)",
                               "Highly Educated (NUTS)", "Unemployment (NUTS)"),
          dep.var.labels = "Attitude Towards the EU (5-point scale)",
          add.lines = list(c("Varying Intercepts", "Yes", "Yes", "Yes", "Yes", "Yes"),
                           c("Varying Slopes", "No", "No", "No", "No", "Yes")))
```

```{r, Variance Components}
summary(ml0)
summary(ml1)
summary(ml2)
summary(ml3)
summary(ml4)

anova(ml1, ml2, ml3, ml4)
```

## Annex A

```{r, Annex A, results='asis'}
stargazer(m.1stage1, m.1stage2, m.1stage3, m.1stage4, type = "html", style = "apsr",
          title = "Annex A. First-Stage IV Results for a Subset of 7 European Countries (2009-2017)",
          dep.var.labels = c("Immigrant Flows (OECD)", "Immigrant Flows (Eurostat)"),
          covariate.labels = c("Shift Share Instrument",
                               "Immigrant Stock",
                               "Unemployment Rate",
                               "Share Aged 25-49",
                               "GDP Growth (in %)",
                               "Globalisation Exposure")
          )
```

## Session Info

```{r}
pander::pander(sessionInfo(), compact = FALSE)
```